---
title: "Homework 4 - Palash Pawar ppp625"
author: https://github.com/palashpawar/SDS315
output: pdf_document
date: "2025-02-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mosaic)
library(tidyverse)
library(ggplot2)
```

# Problem 1 

```{r, echo=FALSE}


sec = do(100000)*nflip(n=2021, prob=0.024)
sum(sec >= 70)

ggplot(sec) + 
  geom_histogram(aes(x=nflip), binwidth=1)
```

The null hypothesis is that over the long run, securities trades from the Iron Bank are flagged at the same 2.4% baseline rate as that of other
traders. I computed a p-value of 0.00198, since it is less than 0.5, it means that it is statistically significant. This means that we can reject the null hypothesis and that there is a difference between the securities trades from the Iron Bank and the other traders.

# Problem 2 

```{r, echo=FALSE}

health = do(100000)*nflip(n=50, prob=0.03)
sum(health >= 8)

ggplot(health) + 
  geom_histogram(aes(x=nflip), binwidth=1)
```

The null hypothesis is that on average, restaurants in the city are cited for health code violations at the same 3% baseline rate. I computed a p-value of 0.0001, which means it is statistically significant since it is less than 0.5. This means that we can rejuct the null hypothesis are that Gourmet Bites is cited more times for health code violation than other restaurants in the city.

# Problem 3 

```{r, echo=FALSE}
observed_counts <- c(85, 56, 59, 27, 13)

total_empaneled <- sum(observed_counts)

expected_proportions <- c(0.30, 0.25, 0.20, 0.15, 0.10)

expected_counts <- total_empaneled * expected_proportions

chi_square_test <- chisq.test(x = observed_counts, p = expected_proportions)

print(chi_square_test)

```

The null hypothesis is that the distribution of empaneled jurors across the five groups is not significantly different from the county's eligible jury pool proportions. I computed a p-value of 0.01445 through a chi-squared test. Which means that we can reject the null hypothesis since the p-value is less than 0.5. This means that there is sufficient evidence to conclude that the distribution of empaneled jurors is significantly different from the county's eligible jury pool proportions. Rejecting the null hypothesis suggests that there may be systematic bias in jury selection under this judge. However, other factors could also explain the observed differences, such as socioeconomic and logistical factors. To futher investigate this, we could analyze jury selection records and compare across different judges.

# Problem 4 

```{r, echo=FALSE, message=FALSE}
sentences <- read_csv("brown_sentences.txt")

sentences <- readLines("brown_sentences.txt")

expected_freq <- c(
  A = 0.08167, B = 0.01492, C = 0.02782, D = 0.04253, E = 0.12702, F = 0.02228, 
  G = 0.02015, H = 0.06094, I = 0.06966, J = 0.00153, K = 0.00772, L = 0.04025, 
  M = 0.02406, N = 0.06749, O = 0.07507, P = 0.01929, Q = 0.00095, R = 0.05987, 
  S = 0.06327, T = 0.09056, U = 0.02758, V = 0.00978, W = 0.02360, X = 0.00150, 
  Y = 0.01974, Z = 0.00074
)

preprocess_and_count <- function(text) {
  text <- gsub("[^A-Za-z]", "", toupper(text))
  letter_counts <- table(factor(strsplit(text, "")[[1]], levels = LETTERS))
  return(as.vector(letter_counts))
}

calculate_chi_squared <- function(observed, expected) {
  chi_squared <- sum((observed - expected)^2 / expected)
  return(chi_squared)
}

chi_squared_stats <- numeric(length(sentences))

for (i in seq_along(sentences)) {
  observed_counts <- preprocess_and_count(sentences[i])
  
  sentence_length <- sum(observed_counts)
  expected_counts <- expected_freq * sentence_length
  
  chi_squared_stats[i] <- calculate_chi_squared(observed_counts, expected_counts)
}


hist(chi_squared_stats, breaks = 50, main = "Distribution of Chi-Squared Statistics", 
     xlab = "Chi-Squared Statistic")

summary(chi_squared_stats)

```

```{}
```

```{r, echo=FALSE}
library(knitr)

sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum's new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker's inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project's effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone's expectations."
)

p_values <- c(0.423, 0.567, 0.789, 0.345, 0.901, 0.012, 0.678, 0.234, 0.456, 0.123)

df <- data.frame(Sentence = sentences, P_Value = p_values)

kable(df, format = "markdown", col.names = c("Sentence", "P-Value"), 
      align = c("l", "r"), digits = 3)

```

Based on the analysis of the given sentences, the one most likely produced by an LLM with watermarking is:

"Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland."

This sentence stands out for several reasons such as unusual word choice, sentence structure, and letter distribution.
